---
title: "Day 3"
author: "Sara"
date: "3/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tibbletime)
```


## Getting the data

For now, read from the local text file
Todo for later to read automatically?

```{r}
data <- read_table("../data/day3.txt", col_names = c("bin_values"))
glimpse(data)
```

## First challenge
separate strings into separate columns
```{r}
data <- data %>%
  separate(bin_values, paste(1:12), sep="(?<=[0-1])")
```

ah! cannot find function to transpose, or function for mode. or i am just too tired.
```{r}
as_tibble(t(data)) %>%
  summarise(mode = mode())
```

## Second challenge
now the calculation gets more complicated: new variable called aim

* down adds to aim
* up removes from aim
* forward adds to horizontal, multiplies with aim then adds to depth

```{r}
heLper_depth <- function(data_tibble) {
  depth <- 0
  aim <- 0
  
  for (i in 1:nrow(data_tibble)){
    my_val <- data_tibble[i, "value"] %>% pull(value)
    if(data_tibble[i, "direction"] == "forward"){
        depth <- depth +my_val  * aim
    } else if (data_tibble[i, "direction"] == "down"){
      aim <- aim + my_val
    }  else if (data_tibble[i, "direction"] == "up"){
      aim <- aim - my_val
    }
  }
  
  return(depth)
}
```

uuhh, just found an ugly false friend between python and R: in python, `length(df)` gives me the number of rows, whereas length in R gives me the number of columns.... whoa.

```{r}
depth <- heLper_depth(data)
hor <- data %>%
  filter(direction == "forward") %>%
  summarise(sum = sum(value)) %>%
  pull(sum)
  
```

```{r}
solution <- hor * depth
solution
```


yay, solved today's puzzle in R as well :-)
