---
title: "Day 1"
author: "Sara"
date: "12/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tibbletime)
```


## Getting the data

For now, read from the local text file
Todo for later to read automatically?

```{r}
data <- read_table("../data/day1.txt", col_names = 'values')
glimpse(data)
```
##First challenge
how many of the differences between subsequent numbers are positive
```{r}
data %>%
  mutate(diff = c(NA,diff(values))) %>%
  filter(diff > 0, na.rm = TRUE) %>%
  nrow()
```

## Second challenge
rather than taking subsequent values, first calculate the sum of a rolling window of 3, and then check how many of those subsequent values are positive

generate appropriate rolling function
```{r}
rolling_sum_of_3 <- rollify(sum, window = 3)
```


```{r}
data %>%
  mutate(rolling_sum = rolling_sum_of_3(values)) %>%
  mutate(diff_rolling = c(NA,diff(rolling_sum))) %>%
  filter(diff_rolling > 0, na.rm = TRUE) %>%
  nrow()
  
```
yay, solved today's puzzle in R as well :-)
