---
title: "Day 2"
author: "Sara"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tibbletime)
```


## Getting the data

For now, read from the local text file
Todo for later to read automatically?

```{r}
data <- read_table("../data/day2.txt", col_names = c("direction", "value"),
                   cols(direction = col_character(),
                                            value = col_integer()))
glimpse(data)
```

## First challenge
sum of forward is x, sum of down minus sum of up is depth, solution is their multiplication
```{r}
my_pivot <- data %>%
  group_by(direction) %>%
  summarise(sum = sum(value))
my_pivot
```

```{r}
x <- my_pivot %>% 
  filter(direction == 'forward') %>%
  pull(sum)
total_down <- my_pivot %>%
  filter(direction == 'down')%>%
  pull(sum)
total_up <- my_pivot %>%
  filter(direction == 'up')%>%
  pull(sum)
depth <- total_down - total_up
solution <- depth * x
solution
```

## Second challenge
now the calculation gets more complicated: new variable called aim

* down adds to aim
* up removes from aim
* forward adds to horizontal, multiplies with aim then adds to depth

```{r}
heLper_depth <- function(data_tibble) {
  depth <- 0
  aim <- 0
  
  for (i in 1:nrow(data_tibble)){
    my_val <- data_tibble[i, "value"] %>% pull(value)
    if(data_tibble[i, "direction"] == "forward"){
        depth <- depth +my_val  * aim
    } else if (data_tibble[i, "direction"] == "down"){
      aim <- aim + my_val
    }  else if (data_tibble[i, "direction"] == "up"){
      aim <- aim - my_val
    }
  }
  
  return(depth)
}
```

uuhh, just found an ugly false friend between python and R: in python, `length(df)` gives me the number of rows, whereas length in R gives me the number of columns.... whoa.

```{r}
depth <- heLper_depth(data)
hor <- data %>%
  filter(direction == "forward") %>%
  summarise(sum = sum(value)) %>%
  pull(sum)
  
```

```{r}
solution <- hor * depth
solution
```


yay, solved today's puzzle in R as well :-)
